Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology

### 2025-05-15 analysis/paper/article.qmd
added more details on duration of editorial activity

# Most editors are only active for less that one year
p_editor_life_distr <-
ggplot(edits_for_each_year_tbl_span) +
  aes(edit_span_years) +
  geom_histogram() +
  theme_minimal(base_size = small_base_size) +
  scale_y_log10() +
  labs(x = "Years editing", y = "Number of editors")

@@

edits_for_each_year_tbl_span_tbl_less_than_one_year <-
edits_for_each_year_tbl_span_tbl %>% 
  mutate(sum_less_than_one = if_else(cumsum(edit_span_years) < 1, n, 0)) %>% pull(sum_less_than_one) %>% sum()

@@

@@

Collaboratively-edited resources such as the Academic Jobs Wiki can be highly variable in the amount and type of editorial activity over time. It is important to characterise this activity to assess the reliability of the content is as a source of information about the job market. The MediaWiki technology used by Academic Jobs Wiki creates a public record of every act of adding or removing text to any page in the Wiki [@barrett2008mediawiki]. @fig-panel-wiki-characteristics, which allows us to assess the dynamics of editing activity over the ten years of our study period. Every edit includes the author's identity, recorded as either an IP address (a numerical address that includes information about the location of the user's computer) or a username. Of the `r n_all_user_ids` unique editors in our sample, `r per_users_we_can_geolocate`% used IP addresses rather than usernames, and geolocation analysis shows that `r results_ip_addresses_tbl_us_scl`% of IP addresses were based in the US, indicating the Archaeology Jobs Wiki is primarily edited and visited by users in the U.S. The number of edits and active editors roughly halved from 2013 to 2017, and then continued to gradually decline. Despite this downward trend, the size of individual edits slightly increases over time, and the number of edits per job ad is relatively constant across our study period. The majority of editors are active on the site for less than a year (n = r`edits_for_each_year_tbl_span_tbl_less_than_one_year`, `r edits_for_each_year_tbl_span_tbl_less_than_one_year_perc`%), perhaps reflecting a single season of job-searching. Editors with a longer history likely come from two groups, job seekers participating in multiple seasons of job applications, and faculty on search committees posting multiple job ads. That said, inferring the number of editors from the number of IP addresses is complicated by the high mobility of job seekers. For example one person could be represented by several IP addresses as they relocate from one city to another while working in a series of short-term appointments such as post-doctoral fellowships and adjunct teaching positions. This might result in an overestimation of the number of editors on the Wiki and underestimation of the duration of their editorial activity.

<!--JB: Ben, this is where I'd like some of the details about min, max, etc. What proportion of people are on the wiki for more than a year, and how long do they stay on? -->

@@

Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-2014. The wiki’s emphasis on tenure-track jobs runs counter to well-established patterns in the academic job market, wherein short-term and contingent positions are fast outpacing “the last good job\[s\] in America” [@aronowitz2001]. Research across anthropological sub-fields has shown that the number of tenure-track positions advertised within the discipline has also declined. Analyzing the job ads posted in *Anthropology News* in 1999–20000 versus 2019–2020, for example, Gershon and Rachok [-@HelloTristesTropes2021] found that the number of tenure-track positions decreased by 25% between their two time slices. Similarly, in his examination of jobs ads posted on the Biological Anthropology Academic Job Wiki between 2010–2017, Passalacqua noted that approximately one quarter of the 474 ads posted were for adjunct or visiting staff [@Passalacqua_2018, pp.773]. Though Passalacqua did not examine how the proportion of job types changed over time, he did underscore “a current diverging trend of decreasing academic job advertisements and increasing doctoral degrees in biological anthropology which could result in serious consequences for the discipline” [-@Passalacqua_2018, pp.773]. Such observations have a surprising longevity. Rogge raised the alarm nearly half a century ago, warning of the “maladaptive” consequences of the exponential rate of disciplinary growth and impending consequences for job seekers [@rogge1976, pp.839]. The dwindling number of permanent academic positions is not a trend unique to anthropology—as of 2019, the AAUP documented a 36% increase in contingent positions over the preceding 15 years (AAUP 2022), and as of 2022, contingent positions make up more than half of faculty positions in the United States [@colby2022annual] .

>>>>

# Most editors are only active for less that one year
p_editor_life_distr <-
  edits_for_each_year_tbl_span %>% 
  mutate(edit_span_years = edit_span_years + 0.001) %>% 
ggplot() +
  aes(edit_span_years) +
  geom_histogram() +
  scale_y_continuous(trans = scales::pseudo_log_trans(sigma = 0.01),
                     breaks = c(0, 1, 10, 100, 1000), minor_breaks = NULL) +
  theme_minimal(base_size = small_base_size) +
  labs(x = "Years editing", y = "Number of editors")

@@

# how many active for less than one day?
edits_for_each_year_tbl_span_tbl_less_than_one_day  <- 
  edits_for_each_year_tbl_span_tbl[1, ]$n

edits_for_each_year_tbl_span_tbl_less_than_one_day_perc  <- 
  round(edits_for_each_year_tbl_span_tbl[1, ]$perc, 0)

# how many active for less than one year?
edits_for_each_year_tbl_span_tbl_less_than_one_year <-
edits_for_each_year_tbl_span_tbl %>% 
  filter(between(edit_span_years, 0.01, 1)) %>% 
  pull(n) %>% sum()

@@

# how many active for more than one year?
edits_for_each_year_tbl_span_tbl_more_than_one_year <-
edits_for_each_year_tbl_span_tbl %>% 
  filter(between(edit_span_years, 1, 10)) %>% 
  pull(n) %>% sum()

edits_for_each_year_tbl_span_tbl_more_than_one_year_perc <- 
  round(edits_for_each_year_tbl_span_tbl_more_than_one_year / sum(edits_for_each_year_tbl_span_tbl$n) * 100, 0)

# average length for those more than one day
edits_for_each_year_tbl_span_tbl_mean <- 
edits_for_each_year_tbl_span_tbl %>% 
  filter(between(edit_span_years, 0.01, 10)) %>% 
  pull(n) %>% mean() %>% round(1)

@@

Collaboratively-edited resources such as the Academic Jobs Wiki can be highly variable in the amount and type of editorial activity over time. It is important to characterise this activity to assess the reliability of the content is as a source of information about the job market. The MediaWiki technology used by Academic Jobs Wiki creates a public record of every act of adding or removing text to any page in the Wiki [@barrett2008mediawiki]. @fig-panel-wiki-characteristics, which allows us to assess the dynamics of editing activity over the ten years of our study period. Every edit includes the author's identity, recorded as either an IP address (a numerical address that includes information about the location of the user's computer) or a username. Of the `r n_all_user_ids` unique editors in our sample, `r per_users_we_can_geolocate`% used IP addresses rather than usernames, and geolocation analysis shows that `r results_ip_addresses_tbl_us_scl`% of IP addresses were based in the US, indicating the Archaeology Jobs Wiki is primarily edited and visited by users in the U.S. The number of edits and active editors roughly halved from 2013 to 2017, and then continued to gradually decline. Despite this downward trend, the size of individual edits slightly increases over time, and the number of edits per job ad is relatively constant across our study period. The majority of editors are active on the site for a single day (n = `r edits_for_each_year_tbl_span_tbl_less_than_one_day`, `r edits_for_each_year_tbl_span_tbl_less_than_one_day_perc`%). There are `r edits_for_each_year_tbl_span_tbl_less_than_one_year` (`r edits_for_each_year_tbl_span_tbl_less_than_one_year_perc`%) editors who were active for between one day and one year. Only `r edits_for_each_year_tbl_span_tbl_more_than_one_year` (n = `r edits_for_each_year_tbl_span_tbl_more_than_one_year_perc`%) editors were active for more than one year, with `r max(round(edits_for_each_year_tbl_span_tbl$edit_span_years,0))` years as the longest period any single editor was active on the Wiki. For those active for more than one day, the average duration of editorial activity is `r edits_for_each_year_tbl_span_tbl_mean` years. This suggests that most of the editing activity on the Wiki reflects a single season of job-searching. Editors with a longer history likely come from two groups, job seekers participating in multiple seasons of job applications, and faculty on search committees posting multiple job ads. That said, inferring the number of editors from the number of IP addresses is complicated by the high mobility of job seekers. For example one person could be represented by several IP addresses as they relocate from one city to another while working in a series of short-term appointments such as post-doctoral fellowships and adjunct teaching positions. This might result in an overestimation of the number of editors on the Wiki and underestimation of the duration of their editorial activity.

@@

Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-2014. The wiki’s emphasis on tenure-track jobs runs counter to well-established patterns in the academic job market, wherein short-term and contingent positions are fast outpacing “the last good job[s] in America” [@aronowitz2001]. Research across anthropological sub-fields has shown that the number of tenure-track positions advertised within the discipline has also declined. Analyzing the job ads posted in *Anthropology News* in 1999–20000 versus 2019–2020, for example, Gershon and Rachok [-@HelloTristesTropes2021] found that the number of tenure-track positions decreased by 25% between their two time slices. Similarly, in his examination of jobs ads posted on the Biological Anthropology Academic Job Wiki between 2010–2017, Passalacqua noted that approximately one quarter of the 474 ads posted were for adjunct or visiting staff [@Passalacqua_2018, pp.773]. Though Passalacqua did not examine how the proportion of job types changed over time, he did underscore “a current diverging trend of decreasing academic job advertisements and increasing doctoral degrees in biological anthropology which could result in serious consequences for the discipline” [-@Passalacqua_2018, pp.773]. Such observations have a surprising longevity. Rogge raised the alarm nearly half a century ago, warning of the “maladaptive” consequences of the exponential rate of disciplinary growth and impending consequences for job seekers [@rogge1976, pp.839]. The dwindling number of permanent academic positions is not a trend unique to anthropology—as of 2019, the AAUP documented a 36% increase in contingent positions over the preceding 15 years (AAUP 2022), and as of 2022, contingent positions make up more than half of faculty positions in the United States [@colby2022annual] .
